opponent modelling

### 2019-11-26 comp_model_1019.Rmd
bug fixes bayes model

## Modelling fixed type participant
```{r}
# FIXED STRATEGIES (level-k) 
# create a function which computes the -2 log Likelihood over the three games for a particular strategy
# modelLL <- function(par,data,strategy, return_value=c("-2logLik","likelihood_by_trial")) {
#   # par = parameter vector (just has one element here) 
#   # dat = data subset for one participant
#   # strategy = c("level1","level2","level3","WSLS")
#   # return = -2logLik (default, for use in optimization) or "likelihood_by_trial" (for later plotting, etc)
#   return_value <- match.arg(return_value)
#   
#   # count the number of supplied parameters. If 1, it is a _1 model, 
#   # if 3, it is a _3 model. For the _3 model we will use unbounded
#   # optimization (transforming parameters from a logit scale back to 0 to 1)
#   n_theta <- length(par)
#   if(n_theta == 3) par <-my_logistic(par)
#   theta <- par[1]
#   # set up a vector to store the likelihood
#   lik <- rep(0.0,nrow(data))
#   for(t in 1:nrow(data)) {
#     t_game <- data[t,"game"]
#     pred_file <- switch(as.character(t_game),rps=rps_predict,fwg = fwg_predict, numbers = numbers_predict)
#     if(n_theta == 3) {
#       theta <- switch(as.character(t_game),rps=par[1],fwg = par[2], numbers = par[3])
#     }
#     if(t_game == "numbers") nopts <- 5 else nopts <- 3
#     if(data[t,"round"] == 1) {
#       # first round is uniform prediction
#       pred <- NA
#       lik[t] <- 1/nopts
#     } else {
#       # get prediction from file
#       pred <- as.character(filter(pred_file,human_previous == as.character(data[t-1,"h_action"]) & computer_previous == as.character(data[t-1,"a_action"]))[strategy][1,])
#       if(as.character(data[t,"h_action"]) == pred) {
#         lik[t] <- theta + (1-theta)/nopts
#       } else {
#         lik[t] <- (1-theta)/nopts
#       }
#     }
#     # for debugging uncomment the line below
#     # cat(as.character(pred),as.character(data[t,"h_action"]),lik[t],"\n")
#   }
#   if(return_value == "-2logLik") {
#     ret <- -2*sum(log(lik))
#     if(ret == Inf) {
#       return(1e+300)
#     } else {
#       return(ret)
#     }
#   }
#   if(return_value == "likelihood_by_trial") return(lik)
# }
# 
# all_modelling <- list()
# for(id in unique(dat$human_id)) {
#   all_modelling[[id]] <- list()
#   tdat <- subset(dat,human_id == id)
#   # level 1
#   all_modelling[[id]]$level1_1 <- optimize(modelLL,interval=c(0,1),data=tdat,strategy="level1")
#   all_modelling[[id]]$level1_3 <- optim(my_logit(rep(all_modelling[[id]]$level1_1$minimum,3)),fn=modelLL,data=tdat,strategy="level1")
#   # level 2
#   all_modelling[[id]]$level2_1 <- optimize(modelLL,interval=c(0,1),data=tdat,strategy="level2")
#   all_modelling[[id]]$level2_3 <- optim(my_logit(rep(all_modelling[[id]]$level2_1$minimum,3)),fn=modelLL,data=tdat,strategy="level2")
#   # level 3
#   all_modelling[[id]]$level3_1 <- optimize(modelLL,interval=c(0,1),data=tdat,strategy="level3")
#   all_modelling[[id]]$level3_3 <- optim(my_logit(rep(all_modelling[[id]]$level3_1$minimum,3)),fn=modelLL,data=tdat,strategy="level3")
#   # WSLS
#   all_modelling[[id]]$WSLS_1 <- optimize(modelLL,interval=c(0,1),data=tdat,strategy="WSLS")
#   all_modelling[[id]]$WSLS_3 <- optim(my_logit(rep(all_modelling[[id]]$WSLS_1$minimum,3)),fn=modelLL,data=tdat,strategy="WSLS")
# }
# 
# save(all_modelling,file="all_modelling.RData")

@@

## Fixed type: Likelihood of the different models by trial

@@

@@

best_response <- function(game,ai_action){
  if (game == "rps"){
    if (ai_action == "R") {return("P")}
    else if (ai_action == "P") {return("S")}
    else if (ai_action == "S") {return("R")}
  } else if (game == "fwg"){ 
    if (ai_action == "F") {return("W")}
    else if (ai_action == "W") {return("G")}
    else if (ai_action == "G") {return("F")}
  } else if (game == "numbers") {
    if (ai_action == "1") {return("2")}
    else if (ai_action == "2") {return("3")}
    else if (ai_action == "3") {return("4")}
    else if (ai_action == "4") {return("5")}
    else if (ai_action == "5") {return("1")}
  }
}
```

```{r}
naive_bayes <- function(par,data,opp_strategy_vec,return_value,opp_mod_transfer=TRUE) {
  # theta = parameter, probability computer will play its best response to what it thinks human is. Truth = 90%. in [0,1].
  # Epsilone: parameter controlling noise in human choice. Probability human will deviate from its best response. In [0,1].
  # dat = data subset for one participant
  # opp_strategy_vec = vector of possible opponent strtegies, model assumes humans restrict opp strategy space to vector   c("level0","level1","level2")
  # return = -2logLik 
  # or "likelihood_by_trial" vector  (for later plotting, etc)
  # opp_mod_transfer is a boolean. True if distribution of opponent strategies is kept across games. False otherwise.

@@

@@

  for(t in 1:nrow(data)) {
    t_game <- data[t,"game"]
    pred_file_opp <- switch(as.character(t_game),rps = rps_predict_opp,fwg = fwg_predict_opp, numbers = numbers_predict_opp)
    if(t_game == "numbers") nopts <- 5 else nopts <- 3
    if(data[t,"round"] == 1) {
      # first round is uniform prediction
      pred <- NA
      lik_hum[t] <- 1/nopts
      if (opp_mod_transfer) {
        opp_prior_vec <- rep(1/num_strat, num_strat)
        lik_opp[,t] <- opp_prior_vec
        }

    # Rounds after first
    } else {
      # get prediction of opponent action from CSV files
      for (strategy in opp_strategy_vec) {
        k = match(strategy,opp_strategy_vec)
        pred_opp <- as.character(filter(pred_file_opp,human_previous == as.character(data[t-1,"h_action"]) & computer_previous == as.character(data[t-1,"a_action"]))[strategy][1,])
        # Given opponent action predicted, what would be human action best response for each opp strat
        br_hum[[k]] <- best_response(t_game,as.character(pred_opp))
        #cat(as.character(br_hum[[k]]), "\n")

        # Multiply prior by likelihood of observation to get posterior. Done here ot take advantage of For loop. 
        if(as.character(data[t,"a_action"]) == pred_opp) {
          lik_opp[k,t] <- (theta + (1-theta)/nopts)*opp_prior_vec[k]
        } else {
          lik_opp[k,t] <- ((1-theta)/nopts)*opp_prior_vec[k]
        }
      }
      # Get which opponent strategy human current actn maps to (if index = 0 then human current action is br to level-0 comp_strat...)
      indices <- which(br_hum %in% as.character(data[t,"h_action"]))
      #cat(as.character(indices),  "\n")

      # THEN likelihood of current human action is just prior on opponent vec (haven't updated priors with curr opp act yet)
      # if action is not predicted by any level-k OM, assume human chooses randomly with prob eps
      if(length(indices) == 0){
        lik_hum[t] <- eps/nopts
      # if same br action for multiple possible ai opponents, add probabilities 
      } else {
        lik_hum[t] <- sum(opp_prior_vec[indices])*(1-eps) + eps/nopts
      }
      # Standardising the probabilities and updating prior
      lik_opp[,t] <- lik_opp[,t]/sum(lik_opp[,t])
      opp_prior_vec <- lik_opp[,t]
    }
    # for debugging uncomment the line below
    #cat(as.character(pred_opp),as.character(data[t,"h_action"]),lik[,t],"\n")
  }
  if(return_value == "-2loglik") {
    ret <- -2*sum(log(lik_hum))
    if(is.infinite(ret) || is.nan(ret)) {
      return(1e+300)
    } else {
      return(ret)
    }
  }
  if(return_value == "likelihood_by_trial") return(lik_hum)
}

@@

# Fitting a bayesiam model WITH TRANSFER of opp strat distribution between games 
TR_bayes_modelling <- list()
for(id in unique(dat$human_id)) {
  TR_bayes_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)
# TR_bayes_modelling[[id]] <- DEoptim(fn=naive_bayes, lower = 0.0, upper = 1.0, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = TRUE, control=list(trace = FALSE,parallelType=1))

TR_bayes_modelling[[id]] <- optim(c(0.1,0.1),fn=naive_bayes,gr = NULL, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = TRUE,lower = c(0.01,0.01), upper = c(0.99,0.99), method="L-BFGS-B")
}

save(TR_bayes_modelling,file="TR_bayes_modelling.RData")

# Fitting a bayesiam model with NO TRANSFER of opp strat distribution between games 
NT_bayes_modelling <- list()
for(id in unique(dat$human_id)) {
  NT_bayes_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)

  # NT_bayes_modelling[[id]] <- DEoptim(fn=naive_bayes, lower = 0.0, upper = 1.0, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = FALSE, control=list(trace = FALSE,parallelType=1))

  NT_bayes_modelling[[id]] <- optim(c(0.1,0.1),fn=naive_bayes,gr = NULL, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = FALSE ,lower = c(0.01,0.01), upper = c(0.99,0.99), method="L-BFGS-B")
}

@@

```{r}
Nbayes_par_results <- data.frame()
for(id in unique(dat$human_id)) {
Nbayes_par_results <- rbind(Nbayes_par_results,
                       data.frame(
                         "id" = id,
                         "condition" = dat[dat$human_id==id,"condition"][1],
                         # BICs 
                         "BIC_transfer" = TR_bayes_modelling[[id]]$value + 2*log(150),
                         "BIC_no_transfer" = NT_bayes_modelling[[id]]$value+ 2*log(150),
                         # Theta is the parameter governing AI stochasticity. Truth is 0.9
                         "theta_transfer" = TR_bayes_modelling[[id]]$par[1],
                         "theta_no_transfer" = NT_bayes_modelling[[id]]$par[1],
                         # Epsilone is parameter showing cases where human deviates from predictions about opponent (lower better)
                         "eps_transfer" = TR_bayes_modelling[[id]]$par[2],
                         "eps_no_transfer" = NT_bayes_modelling[[id]]$par[2]))
}

@@

## Q-learning
```{r}
Q_learn <- function(par,data,return_value=c("-2loglik","likelihood_by_trial")){
  # actions
  # beta= inverse temperature parameter in softmax choice function
  # lambda = learning rate (one per game?)
  # data : participant and ai choice data.
  # Returns Q-values per trial and predicts choice using softmax
  beta <- par[1]
  alpha <- par[2]
  Q_vals_RPS = matrix(0.0,3)
  names(Q_vals_RPS) <- c("R","P","S")
  Q_vals_FWG = matrix(0.0,3)
  names(Q_vals_FWG) <- c("F","W","G")
  Q_vals_NUM = matrix(0.0,5)
  names(Q_vals_NUM) <- c("1","2","3","4","5")

@@

##Q-learning fit
```{r}
QL_modelling <- list()
for(id in unique(dat$human_id)) {
  QL_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)
  QL_modelling[[id]] <- optim(c(1,0.1),fn=Q_learn,gr = NULL, data=tdat,"-2loglik", lower = c(0,0), upper = c(10,0.99), method="L-BFGS-B")
}
save(QL_modelling,file="QL_modelling.RData")

@@

##Q-Learning with state space consisting of last round play
```{r}

@@

    } else {
      # Get past human action and associated reward 
      h_act <- as.character(data[t-1,"h_action"])
      ai_act <- as.character(data[t-1,"a_action"])
      reward <- as.numeric(data[t-1,"score"])
    }
    # cat(curr_state, " This is the previous state", "\n")
    new_state <- paste0(h_act,ai_act)
    # cat(new_state,"This is new state","\n")

@@

                         # Q-learning with last round states
                         "QL_states_BIC" = QL_states_modelling[[id]]$value + 2*log(150),
                         # beta ->  inverse temperature parameter in softmax choice function
                         "QL_states_Beta" = QL_states_modelling[[id]]$par[1],
                         # alpha -> learning rate in QL update
                         "QL_states_alpha" = QL_states_modelling[[id]]$par[2],

@@

Corr_Tr_model <- data.frame()
for(id in unique(dat$human_id)) {
  tdat <- subset(dat,human_id == id)
  tot_score <- sum(tdat$score)
  tot_time <- sum(tdat$human_rt)
  early_dat <- subset(tdat,between(tdat$round,2,6) & (game =="fwg" | game =="numbers") )
  tr_score <- sum(early_dat$score)
  id_results <- subset(All_results, ID == id)
  min_BIC <- apply(id_results[,c("Random_BIC","Bayes_Tr_BIC","Bayes_No_Tr_BIC","QL_BIC","QL_states_BIC","EWA_BIC","EWA_self_BIC")],1,min)

  best_model <- c("random","Bayes Tr","Bayes No Tr","QL", "QL_states","EWA","S_EWA")[apply(id_results[,c("Random_BIC","Bayes_Tr_BIC","Bayes_No_Tr_BIC","QL_BIC","QL_states_BIC","EWA_BIC","EWA_self_BIC")],1,which.min)]
  # 
  Corr_Tr_model <- rbind(Corr_Tr_model ,
                       data.frame(
                         "id" = id,
                         "condition" = dat[dat$human_id==id,"condition"][1],
                         "Total_time" = sum(tdat$human_rt),
                         "BIC_diff_Bayes" = id_results[,"Bayes_Tr_BIC"] - id_results[,"Bayes_No_Tr_BIC"],
                         "BIC_diff_rand" =  id_results[,"Random_BIC"] - min_BIC,
                         "Early_game_score" = tr_score,
                         "Total_score" = tot_score,
                         "Best_model" = best_model
                       ))
}

# Correlate difference between BICs of Bayes transfer and no transfer with early rounds score (evidence for transfer) 
cor.test(Corr_Tr_model$BIC_diff_Bayes,Corr_Tr_model$Early_game_score, method=c("pearson", "kendall", "spearman"))

# Correlation between early game score and difference between best model and random BIC
cor.test(Corr_Tr_model$BIC_diff_rand, Corr_Tr_model$Early_game_score, method=c("pearson", "kendall", "spearman"))

```
# Notes : There is positive and stat significant correlation between the difference between the best model BIC and random BIC on one hand, and the early scores from the target games ( FWG and NUM).

@@

Corr_Tr_model$model <- recode(Corr_Tr_model$Best_model,"Bayes Tr" = "Tranfer","Bayes No Tr" = "No Transfer", .default = "Random")

model <- factor(Corr_Tr_model$model)
# Total score by best predictive model 
tapply(Corr_Tr_model$Total_score, model, mean)

# Compute the analysis of variance
res.aov <- aov(Total_score ~ model, data = Corr_Tr_model)
# Summary of the analysis
summary(res.aov)
TukeyHSD(res.aov)

## What if we remove from data people who play randomly and then compare Transfer v No Transfer? 
Bayes_only <- subset(Corr_Tr_model, model != "Random" )

Bayes.aov <- aov(Early_game_score ~ model, data = Bayes_only)
# Summary of the analysis
summary(Bayes.aov)
TukeyHSD(Bayes.aov)

wilcox.test(Early_game_score ~ model,data = Bayes_only)

# Total time by best predictive model
tapply(Corr_Tr_model$Total_time, model, mean)
time.aov <- aov(Total_time ~ model, data = Corr_Tr_model)
# Summary of the analysis
summary(time.aov)
TukeyHSD(time.aov)

``` 
#Scores are indeed higher but difference is not statistically significant.

>>>>

## Modelling fixed level-k type participant
```{r}
# FIXED STRATEGIES (level-k) 
# create a function which computes the -2 log Likelihood over the three games for a particular strategy
# modelLL <- function(par,data,strategy, return_value=c("-2logLik","likelihood_by_trial")) {
#   # par = parameter vector (just has one element here) 
#   # dat = data subset for one participant
#   # strategy = c("level1","level2","level3","WSLS")
#   # return = -2logLik (default, for use in optimization) or "likelihood_by_trial" (for later plotting, etc)
#   return_value <- match.arg(return_value)
#   
#   # count the number of supplied parameters. If 1, it is a _1 model, 
#   # if 3, it is a _3 model. For the _3 model we will use unbounded
#   # optimization (transforming parameters from a logit scale back to 0 to 1)
#   n_theta <- length(par)
#   if(n_theta == 3) par <-my_logistic(par)
#   theta <- par[1]
#   # set up a vector to store the likelihood
#   lik <- rep(0.0,nrow(data))
#   for(t in 1:nrow(data)) {
#     t_game <- data[t,"game"]
#     pred_file <- switch(as.character(t_game),rps=rps_predict,fwg = fwg_predict, numbers = numbers_predict)
#     if(n_theta == 3) {
#       theta <- switch(as.character(t_game),rps=par[1],fwg = par[2], numbers = par[3])
#     }
#     if(t_game == "numbers") nopts <- 5 else nopts <- 3
#     if(data[t,"round"] == 1) {
#       # first round is uniform prediction
#       pred <- NA
#       lik[t] <- 1/nopts
#     } else {
#       # get prediction from file
#       pred <- as.character(filter(pred_file,human_previous == as.character(data[t-1,"h_action"]) & computer_previous == as.character(data[t-1,"a_action"]))[strategy][1,])
#       if(as.character(data[t,"h_action"]) == pred) {
#         lik[t] <- theta + (1-theta)/nopts
#       } else {
#         lik[t] <- (1-theta)/nopts
#       }
#     }
#     # for debugging uncomment the line below
#     # cat(as.character(pred),as.character(data[t,"h_action"]),lik[t],"\n")
#   }
#   if(return_value == "-2logLik") {
#     ret <- -2*sum(log(lik))
#     if(ret == Inf) {
#       return(1e+300)
#     } else {
#       return(ret)
#     }
#   }
#   if(return_value == "likelihood_by_trial") return(lik)
# }
# 
# all_modelling <- list()
# for(id in unique(dat$human_id)) {
#   all_modelling[[id]] <- list()
#   tdat <- subset(dat,human_id == id)
#   # level 1
#   all_modelling[[id]]$level1_1 <- optimize(modelLL,interval=c(0,1),data=tdat,strategy="level1")
#   all_modelling[[id]]$level1_3 <- optim(my_logit(rep(all_modelling[[id]]$level1_1$minimum,3)),fn=modelLL,data=tdat,strategy="level1")
#   # level 2
#   all_modelling[[id]]$level2_1 <- optimize(modelLL,interval=c(0,1),data=tdat,strategy="level2")
#   all_modelling[[id]]$level2_3 <- optim(my_logit(rep(all_modelling[[id]]$level2_1$minimum,3)),fn=modelLL,data=tdat,strategy="level2")
#   # level 3
#   all_modelling[[id]]$level3_1 <- optimize(modelLL,interval=c(0,1),data=tdat,strategy="level3")
#   all_modelling[[id]]$level3_3 <- optim(my_logit(rep(all_modelling[[id]]$level3_1$minimum,3)),fn=modelLL,data=tdat,strategy="level3")
#   # WSLS
#   all_modelling[[id]]$WSLS_1 <- optimize(modelLL,interval=c(0,1),data=tdat,strategy="WSLS")
#   all_modelling[[id]]$WSLS_3 <- optim(my_logit(rep(all_modelling[[id]]$WSLS_1$minimum,3)),fn=modelLL,data=tdat,strategy="WSLS")
# }
# 
# save(all_modelling,file="all_modelling.RData")

@@

## Fixed level-k type: Likelihood of the different models by trial

@@

#### New work

@@

#Fit a bayesian model to observed human play. Prior is vector of possible level-k opponent strategies.
```{r}
# Function to produce best reponse to ai_action 
best_response <- function(game,ai_action){
  if (game == "rps"){
    if (ai_action == "R") {return("P")}
    else if (ai_action == "P") {return("S")}
    else if (ai_action == "S") {return("R")}
  } else if (game == "fwg"){ 
    if (ai_action == "F") {return("W")}
    else if (ai_action == "W") {return("G")}
    else if (ai_action == "G") {return("F")}
  } else if (game == "numbers") {
    if (ai_action == "1") {return("2")}
    else if (ai_action == "2") {return("3")}
    else if (ai_action == "3") {return("4")}
    else if (ai_action == "4") {return("5")}
    else if (ai_action == "5") {return("1")}
  }
}

naive_bayes <- function(par,data,opp_strategy_vec,return_value,opp_mod_transfer) {
  # theta = parameter, probability computer will play its best response to what it thinks human is. Truth = 90%. in [0,1].
  # Epsilone: parameter controlling noise in human choice. Probability human will deviate from its best response. In [0,1].
  # dat = data subset for one participant
  # opp_strategy_vec = vector of possible opponent strtegies, model assumes humans restrict opp strategy space to vector   c("level0","level1","level2")
  # return = -2logLik 
  # or "likelihood_by_trial" vector  (for later plotting, etc)
  # opp_mod_transfer is a boolean. True if distribution of opponent strategies is kept across games. False otherwise.

@@

  # Load tables predicting oppponent move
  rps_predict_opp <- read.csv("rps_predict_opp.csv")
  fwg_predict_opp <- read.csv("fwg_predict_opp.csv")
  numbers_predict_opp <- read.csv("numbers_predict_opp.csv")

@@

  for(t in 1:nrow(data)) {
    t_game <- data[t,"game"]
    pred_file_opp <- switch(as.character(t_game),rps = rps_predict_opp,fwg = fwg_predict_opp, numbers = numbers_predict_opp)

    if(t_game == "numbers") nopts <- 5 else nopts <- 3
    if(data[t,"round"] == 1) {
      # first round is uniform prediction
      pred_opp <- NA
      lik_hum[t] <- 1/nopts
      if (!(opp_mod_transfer)) {
        opp_prior_vec <- rep(1/num_strat, num_strat)
        lik_opp[,t] <- opp_prior_vec
        }

    # Rounds after first
    } else {
      # get prediction of opponent action from CSV files
      for (strategy in opp_strategy_vec) {
        k = match(strategy,opp_strategy_vec)
        pred_opp <- as.character(filter(pred_file_opp, pred_file_opp$human_previous == as.character(data[t-1,"h_action"]) & pred_file_opp$computer_previous == as.character(data[t-1,"a_action"]))[[k]])

        # Given opponent action predicted, what would be human action best response for each opp strat
        br_hum[[k]] <- best_response(t_game,as.character(pred_opp))

        # Multiply prior by likelihood of observation to get posterior. Done here ot take advantage of For loop. 
        if(as.character(data[t,"a_action"]) == pred_opp) {
          lik_opp[k,t] <- (theta + (1-theta)/nopts)*opp_prior_vec[k]
        } else {
          lik_opp[k,t] <- ((1-theta)/nopts)*opp_prior_vec[k]
        }
      }
      # Get which opponent strategy human current actn maps to (if index = 0 then human current action is br to level-0 comp_strat...)
      indices <- which(br_hum %in% as.character(data[t,"h_action"]))

      # THEN likelihood of current human action is just prior on opponent vec (haven't updated priors with curr opp act yet)
      # if action is not predicted by any level-k OM, assume human chooses randomly with prob eps
      if(length(indices) == 0){
        lik_hum[t] <- eps/nopts
      # if same br action for multiple possible ai opponents, add probabilities 
      } else {
        lik_hum[t] <- sum(opp_prior_vec[indices])*(1-eps) + eps/nopts
      }
      # Standardising the probabilities and updating prior
      lik_opp[,t] <- lik_opp[,t]/sum(lik_opp[,t])
      opp_prior_vec <- lik_opp[,t]
    }
    # for debugging uncomment the line below
    #cat(as.character(pred_opp),as.character(data[t,"h_action"]),lik[,t],"\n")
  }
  if(return_value == "-2loglik") {
    ret <- -2*sum(log(lik_hum))
    if(is.infinite(ret) || is.nan(ret)) {
      return(1e+300)
    } else {
      return(ret)
    }
  }
  #if(return_value == "likelihood_by_trial") return(lik_hum)
}

@@

# Fitting a bayesiam model WITH TRANSFER of opp strat distribution between games 
TR_bayes_modelling <- list()
for(id in unique(dat$human_id)) {
  TR_bayes_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)

  TR_bayes_modelling[[id]] <- DEoptim(fn=naive_bayes, lower = c(0.0,0.0), upper = c(1.0,1.0), data=tdat, opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = TRUE, control=list(trace = FALSE, parallelType=0,parVar = c("best_response")))

# TR_bayes_modelling[[id]] <- optim(c(0.1,0.1),fn=naive_bayes,gr = NULL, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = TRUE,lower = c(0.01,0.01), upper = c(0.99,0.99), method="L-BFGS-B")
}
save(TR_bayes_modelling,file="TR_bayes_modelling.RData")

# Fitting a bayesian model with NO TRANSFER of opp strat distribution between games 
```{r}
NT_bayes_modelling <- list()
for(id in unique(dat$human_id)) {
  NT_bayes_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)

  NT_bayes_modelling[[id]] <- DEoptim(fn=naive_bayes, lower = c(0.0,0.0), upper = c(1.0,1.0), data=tdat, opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = FALSE, control=list(trace = FALSE, parallelType=0,parVar = c("best_response")))

#   NT_bayes_modelling[[id]] <- optim(c(0.1,0.1),fn=naive_bayes,gr = NULL, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = FALSE ,lower = c(0.01,0.01), upper = c(0.99,0.99), method="L-BFGS-B")
}

@@

# Summarise results in dataframe for both transfer and no transfer models 
```{r}
Nbayes_par_results <- data.frame()
for(id in unique(dat$human_id)) {
Nbayes_par_results <- rbind(Nbayes_par_results,
                       data.frame(
                         "id" = id,
                         "condition" = dat[dat$human_id==id,"condition"][1],
                         # BICs 
                         "BIC_transfer" = TR_bayes_modelling[[id]]$value + 2*log(150),
                         "BIC_no_transfer" = NT_bayes_modelling[[id]]$value+ 2*log(150),
                         # Theta is the parameter governing AI stochasticity. Truth is 0.9
                         "theta_transfer" = TR_bayes_modelling[[id]]$par[1],
                         "theta_no_transfer" = NT_bayes_modelling[[id]]$par[1],
                         # Epsilone is parameter showing cases where human deviates from predictions about opponent (lower better)
                         "eps_transfer" = TR_bayes_modelling[[id]]$par[2],
                         "eps_no_transfer" = NT_bayes_modelling[[id]]$par[2]))
}

@@

## Q-learning (Basic)
```{r}
Q_learn <- function(par,data,return_value=c("-2loglik","likelihood_by_trial")){
  # actions
  # beta= inverse temperature parameter in softmax choice function
  # lambda = learning rate (one per game?)
  # data : participant and ai choice data.
  # Returns Q-values per trial and predicts choice using softmax
  beta <- par[1]
  alpha <- par[2]
  Q_vals_RPS = matrix(0.0,3)
  names(Q_vals_RPS) <- c("R","P","S")
  Q_vals_FWG = matrix(0.0,3)
  names(Q_vals_FWG) <- c("F","W","G")
  Q_vals_NUM = matrix(0.0,5)
  names(Q_vals_NUM) <- c("1","2","3","4","5")

@@

## Q-learning fit
```{r}
QL_modelling <- list()
for(id in unique(dat$human_id)) {
  QL_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)
  QL_modelling[[id]] <- optim(c(1,0.1),fn=Q_learn,gr = NULL, data=tdat,"-2loglik", lower = c(0,0), upper = c(10,0.99), method="L-BFGS-B")
}
save(QL_modelling,file="QL_modelling.RData")

@@

## Q-Learning with state space consisting of last round play
```{r}

@@

    } else {
      # Get past human action and associated reward 
      h_act <- as.character(data[t-1,"h_action"])
      ai_act <- as.character(data[t-1,"a_action"])
      reward <- as.numeric(data[t-1,"score"])
    }

    # cat(curr_state, " This is the previous state", "\n")
    new_state <- paste0(h_act,ai_act)
    # cat(new_state,"This is new state","\n")

@@

                         # Q-learning with last round states
                         "QL_states_BIC" = QL_states_modelling[[id]]$optim$bestval + 2*log(150),
                         # beta ->  inverse temperature parameter in softmax choice function
                         "QL_states_Beta" = QL_states_modelling[[id]]$optim$bestmem[1],
                         # alpha -> learning rate in QL update
                         "QL_states_alpha" = QL_states_modelling[[id]]$optim$bestmem[2],

@@

Corr_Tr_model <- data.frame()
for(id in unique(dat$human_id)) {
  tdat <- subset(dat,human_id == id)
  tot_score <- sum(tdat$score)
  tot_time <- sum(tdat$human_rt)
  early_dat <- subset(tdat,between(tdat$round,2,11) & (game =="fwg" | game =="numbers") )
  tr_score <- sum(early_dat$score)
  id_results <- subset(All_results, ID == id)
  min_BIC <- apply(id_results[,c("Random_BIC","Bayes_Tr_BIC","Bayes_No_Tr_BIC","QL_BIC","QL_states_BIC","EWA_BIC","EWA_self_BIC")],1,min)

  best_model <- c("random","Bayes Tr","Bayes No Tr","QL", "QL_states","EWA","S_EWA")[apply(id_results[,c("Random_BIC","Bayes_Tr_BIC","Bayes_No_Tr_BIC","QL_BIC","QL_states_BIC","EWA_BIC","EWA_self_BIC")],1,which.min)]
  # 
  Corr_Tr_model <- rbind(Corr_Tr_model ,
                       data.frame(
                         "id" = id,
                         "condition" = dat[dat$human_id==id,"condition"][1],
                         "Early_game_score" = tr_score,
                         "Total_score" = tot_score,
                         "Best_model" = best_model,
                         "Total_time" = sum(tdat$human_rt),
                         "TR_minus_NT_BIC" = id_results[,"Bayes_Tr_BIC"] - id_results[,"Bayes_No_Tr_BIC"],
                         "Rand_minus_best_BIC" =  id_results[,"Random_BIC"] - min_BIC

                       ))
}
```

# Correlate difference between BICs of Bayes transfer and no transfer with early rounds score (evidence for transfer) 
```{r}

cor.test(Corr_Tr_model$TR_minus_NT_BIC,Corr_Tr_model$Early_game_score, method="spearman")

#Correlation between early game score and difference between best model and random BIC
```{r}
cor.test(Corr_Tr_model$Rand_minus_best_BIC, Corr_Tr_model$Early_game_score, method="spearman")

```
# Notes : There is positive and stat significant correlation between the difference between the best model BIC and random BIC on one hand, and the early scores from the target games ( FWG and NUM). What does this mean? It can be interpreted as a validation of early game score as good measure of how non-random play is....

@@

Corr_Tr_model$model <- recode(Corr_Tr_model$Best_model,"Bayes Tr" = "Tranfer","Bayes No Tr" = "No Transfer", .default = "Random")
model <- factor(Corr_Tr_model$model)
condition <- factor(Corr_Tr_model$condition)

# Total score by best predictive model 
tapply(Corr_Tr_model$Total_score, model, mean)

# Compute the analysis of variance
res.aov <- aov(Total_score ~ model + condition, data = Corr_Tr_model)
# Summary of the analysis
summary(res.aov)
TukeyHSD(res.aov)
```

## What if we remove from data people who play randomly and then compare Transfer v No Transfer? 
```{r}

Bayes_only <- subset(Corr_Tr_model, model != "Random" )

group_by(Bayes_only, model) %>%
  summarise(
    count = n(),
    mean = mean(Early_game_score, na.rm = TRUE),
    sd = sd(Early_game_score, na.rm = TRUE)
  )

wilcox.test(Early_game_score ~ model, data = Bayes_only) 
```

# Total time by best predictive model
```{r}
tapply(Corr_Tr_model$Total_time, model, mean)

wilcox.test(Early_game_score ~ model, data = Bayes_only)

time.aov <- aov(Total_time ~ model, data = Corr_Tr_model)
# Summary of the analysis
summary(time.aov)
TukeyHSD(time.aov)

``` 
#Early game scores for people best modelled with a Bayesian model with transfer are indeed higher than those best modelled without transfer but difference is not statistically significant.
